Bandwidth-Aware Reconfigurable Memory Design with Hybrid Memory Technologies

### 2011-04-05 paper.tex

Based on our memory technology exploration, we now provide an overview of BROM
design. While many design methods envolved with new memory technologies endeavor
to reduce the off-chip memory access latency, this work focuses on minimize
off-chip bandwidth requirement by employing hybrid on-chip memory and
reconfiguration. For each level of memory, we select the memory technology that
has the highest provided bandwidth within given capacity. On top of such
hardware hierarchy design, we apply dynamic memory system reconfiguration with
the bandwidth demand of a specific set of applications as the major metric. One
desirable characteristic of BROM is that it provides prioritized access to the
residual memory capacity available at runtime instead of sharing it among all
data flows. By having control over the residual capacity, we can better utilize
the limited resources to enhance system performance. The hardware
reconfiguration is further facilitated by software prediction engine. Rather
than conventional last value predictors and table-based predictors, we employ a
probablity-based predictor that can achieve higher accuracy with small
performance overhead.

\subsection{Memory Hierarchy Design}

@@

tailoring mode
reuse mode

@@

statistical prediction
performance opt

>>>>

Based on our memory technology exploration, we now provide an overview of BAROM
design. While many design methods envolved with new memory technologies endeavor
to reduce the off-chip memory access latency, this work focuses on minimize
off-chip bandwidth requirement by employing hybrid on-chip memory and
reconfiguration. For each level of memory, we select the memory technology that
has the highest provided bandwidth within given capacity. On top of such
hardware hierarchy design, we apply dynamic memory system reconfiguration with
the bandwidth demand of a specific set of applications as the major metric. One
desirable characteristic of BAROM is that it provides prioritized access to the
residual memory capacity available at runtime instead of sharing it among all
data flows. By having control over the residual capacity, we can better utilize
the limited resources to enhance system performance. The hardware
reconfiguration is further facilitated by software prediction engine. Rather
than conventional last value predictors and table-based predictors, we employ a
probablity-based predictor that can achieve higher accuracy with small
performance overhead. The rest of this section describes the BAROM design, which
is composed of on-chip memory hierarchy hardware design, the reconfiguration
mechanism, and the predictor engine.

\subsection{Hardware Configuration}

Figure~\ref{fig:hw-config} shows an overview of BAROM hardware configuration.
The CMP system consists of multiple cores, where L1 cahces are private to each
core and lower level caches are shared by the cores. With the provided bandwidth
vs. capacity curves of various memory technologies that are obtained using the
method described in section~\ref{sec:tech-explore}, we can design a hybrid
on-chip memory system to always provide the highest bandwith. In order to
achieve this goal, we configure the memory hierarchy in the following
manner.\vspace{5pt}

\noindent{}{\bf{Number of levels.}} In order to decide the number of memory
levels, we examine the provided bandwidth curve. Memory levels are seperated by
the cross points of two memory technologies. For example,
Figure~\ref{fig:memory-bw} illustrates that there are two cross points (SRAM and
STT-RAM, STT-RAM and RRAM). Therefore, we will have three levels of shared
caches.\vspace{5pt}

\noindent{}{\bf{Memory technology of each level.}} The memory technology with
the highest provided bandwidth within the given range of capacity of each level
is selected. In our case, SRAM, STT-RAM, and RRAM are selected as the L2, L3,
and L4 caches respectively. \vspace{5pt}

\noindent{}{\bf{Capacity of each level.}} The maximum availabe capacity of each
level is determined by the cross point of two memory technologies. In our case,
the capacities of SRAM/L2, STT-RAM/L2, and RRAM/L4 are 2MB, 512MB, and 1GB
respectively. At the system initialization, the capacity of each level is
defined to be the maximum available capacity of the level. According to the
reqirement of different applications, the capacities can be reconfigured at
run-time to provide best performance. The memory space of each level can also be
reconfigured to form an internal level as shown in Figure~\ref{fig:hw-config}.
The hardware and software supports for reconfiguration will be presented in the
following sections.\vspace{5pt}

The goal of such architecture design is to make use of different memory
technologies to configure an on-chip memory system with optimal provided
bandwidth over a range of capacities. As a result, the overall provided
bandwidth curve of the on-chip memory system will appear to be an envelope line
of all the available memory technologies as illustrated in
Figure~\ref{fig:pbw-envelope}. It is illustrated in this figure that the hybrid
memory system always provides the highest bandwidth over different capacities.

@@

Although the above hybrid memory configuration maintains the optimal provided
bandwidth curve over all the capacities, it does not garantee the best
performance for different applications with a variety of bandwidth requirements.
The overall provided bandwidth curve shown in Figure~\ref{fig:pbs-envelope}
appears to be monotonically decreasing. A specific application may have a
bandwidth requirement that can be mapped to a single point on the curve. In
order to satisfy the bandwidth demand of different applications, we dynamically
reconfigure the memory space of each level corresponding to the bandwidth
requirement of each set of applications running on the system. To be more
aggressive, the reconfiguration is applied at the end of each given execution
time interval to accomodate different program phases of applications.

The major system parameter to be reconfigured is the capacity of each memory
level. At each time interval, we determine the upper bound of the capacity of
memory level-i ($s_{iu}$) by mapping the demand bandwidth (DBW) of a specific
application to the system's provided bandwidth curve, i.e.,
$s_{iu}=f^{-1}(\textrm{DBW})$ where $f(x)$ represents the provided bandwidth
curve of the on-chip memory system stored in the look-up-table in a reserved
non-volatile memory space. The demand bandwidth for each memory level is
generated using the prediction engine presented in section~\ref{sec:predict}. In
previous sections, we show that the memory with smaller capacity provides higher
bandwidth. Theoritically, the higher bandwidth provided by the memory system
leads to higher performance (in terms of throughput) and power consumption of
throughput computing applications~\cite{throughput-vs-bandwidth}. Consequently,
we define a lower bound of memory capacity as
$s_{il}=f^{-1}(\textrm{DBW}*(1+\sigma))$, where $\sigma$ is a pre-defined
threshold to make sure possible provided bandwidth increase without much power
overhead. The capacity of memory level-i ($s_i$) is therefore selected to
satisfy $s_{il}\le s_i\le s_{iu}$.

Figure~\ref{fig:reconfig} describes the hardware support for reconfiguration.

reconfiguration support: how reconfigurable (Li -> Li+1 (Li+0.5); hardware
modification to support reconfiguration.

reconfigure the unused capacity:

- if (> si): becomes si+1

- else: unused (or becomes lut)

@@

statistical prediction

hierarchical prediction

performance opt: lut long enough but not too long (num of entries, bits per
entry)
